---
title: "PA1_template"
author: "Liang Weikang"
output: html_document
---
##1.Loading and preprocessing the data

```{r}
setwd("E:\\My stuffs\\Self Study\\Analytics courses\\coursera\\reproducible research\\assg 1")
library("data.table")
library("plyr")
library(ggplot2)
library(lattice)
activityData = data.table(read.csv("activity.csv",header=TRUE,na.strings="NA", sep=","))
activityData$date = as.Date(activityData$date,format='%Y-%m-%d')

```

##2.What is mean total number of steps taken per day?

```{r}
result=as.data.frame(aggregate(activityData$steps,list(activityData$date),sum))
colnames(result)=c("date","totalSteps")
```

###Histogram of total Steps per day
```{r}
diagram=ggplot(result,aes(x=totalSteps))
diagram+geom_histogram()
```

###Mean total number of steps per day
```{r}
meanTotalSteps = as.numeric(mean(result$totalSteps,na.rm=TRUE))
print(meanTotalSteps)
```

###Median total number of steps per day
```{r}
meanMedianSteps = as.numeric(median(result$totalSteps,na.rm=TRUE))
print(meanMedianSteps)
```


##3.What is the average daily activity pattern?
```{r}
#Time series plot of average steps across all days
ts= as.data.frame(activityData[,j=list(mean(steps,na.rm=TRUE)),by=interval])
colnames(ts) = c("interval","avgStepsDaily")
line = ggplot(ts,aes(x=interval,y=avgStepsDaily))
line + geom_line()+scale_x_continuous("interval",breaks=c(0,800,1200,1600,2000)) + scale_y_continuous("Avg Steps Daily")
```

###Which 5-minute interval, on average across all the days in the dataset, 
###contains the maximum number of steps?
```{r}
ts[head(order(ts$avgStepsDaily,decreasing=TRUE),n=1),]
```


##4.Imputing missing values
```{r}
#total number of NA Values in steps:
table(is.na(activityData$steps))[[2]]
#impute NAs with mean for that 5 min interval
#1. join with original data by interval column to map avgSteps values to other columns
imputedData=join(activityData,ts,by="interval")
#2. converts steps to double value since we are equating it to avgStepsDaily, a double data type
imputedData$steps=as.double(imputedData$steps)
imputedData[is.na(steps),]$steps = imputedData[is.na(steps),]$avgStepsDaily
```

###Histogram of total Steps per day
```{r}
imputedResult=as.data.frame(imputedData[,sum(steps),by="date"])
colnames(imputedResult)=c("date","totalSteps")
diagram2=ggplot(imputedResult,aes(x=totalSteps))
diagram2+geom_histogram()
```
###mean of total steps per day (imputed)
```{r}
mean(imputedResult$totalSteps)
```
###median of total steps per day (imputed)
```{r}
median(imputedResult$totalSteps)
```
####conclusion: imputing has little impact on total steps estimation


##5.Are there differences in activity patterns between weekdays and weekends?
```{r}
#1. determine weekType (weekday or weekend)
imputedData$dayOfWeek=weekdays(imputedData$date)
imputedData$isWeekend=imputedData$dayOfWeek == "Saturday" | imputedData$dayOfWeek == "Sunday"
imputedData$weekType= as.factor(ifelse(imputedData$isWeekend,"weekend","weekday"))

#2. compute avg steps by interval and weektype via aggregation
chartData = imputedData[,mean(steps),by=c("interval","weekType")]
setnames(chartData,"V1","avgSteps")
#3. plot graph separated by weekType
xyplot(avgSteps~interval|weekType,data=chartData,
       type="l",grid=TRUE,layout=c(1,2),
       xlab="Interval",
       ylab="Number of steps"
      )
wkendData= subset(chartData,weekType=="weekend")
wkDayData= subset(chartData,weekType="weekday")
```
###max no. of steps on weekdays
```{r}
wkDayData[head(order(wkDayData$avgSteps,decreasing=TRUE),n=1),]
```
###max no. of steps on weekends
```{r}
wkendData[head(order(wkendData$avgSteps,decreasing=TRUE),n=1),]
```
####conclusion: The xy plot shows major differences in walking patterns the entire day. People tend to be more active throughout the day (from 1000 to 2000 hours) on weekends, as compared to the lesser number of average steps taken on weekdays. The maximum number of average steps on weekdays is also lesser than on weekends.
